---
title: "Regression"
author: "Ahmad Mirzanezhad"
date: "2023-01-13"
output: html_document
---
  
```{r,warning = FALSE, message = FALSE}
require(dplyr)
require(tidyverse)
require(ggplot2)
require(datasets)
require(readr)
library(caret)
library(splines)
library(splines2)
library(mgcv)
library(vctrs)
library(gratia)
df = read.csv("D:/EXETER/Exams-20230720T014746Z-001/Exams/MTHM503/weight_height.csv")
```

To begin with, we start with a simple graph


```{r}

    df %>% 
  pivot_longer(-c(weight, male)) %>%
  ggplot(aes(x = value, y = weight, color=factor(male))) + 
  geom_point() + 
  geom_smooth(method = 'lm', formula = y ~ x ,se= FALSE)+
  guides(color = guide_legend(title = "Male"))+
  facet_wrap(~name, scales = 'free')
       

  
```

the graph obviously shows that the relationship between variables and "weight" is not
linear. Also, there is a difference between male and female in "weight" VS height and age.
So I will not use linear regression to predict the model. In this case, two column will be added to our "dataframe" to see the effect of gender.
## first attempt: 

```{r}
df_sex_effect = mutate(df , male_height = male* height, male_age = male * age) 
model_sex_effect <- lm(weight ~ poly(height, 2, raw=TRUE)+poly(age, 2, raw = TRUE)+ male_height + male_age,data = df_sex_effect)
summary(model_sex_effect)
```
Residual

```{r}
ggplot(model_sex_effect, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0)+
  geom_smooth()+
  labs(title='Residual vs. Fitted Values Plot', x='Fitted Values', y='Residuals')

```

There is a pattern in residual plot. On the other hand, two new variables are not 
significant. Therefor,we will not continue with those two new variables in our model. 

```{r}
set.seed(123)
sample_id <- df$weight %>% 
  createDataPartition(p = 0.8, list = FALSE)
train_data <- df[sample_id, ]
test_data <- df[-sample_id, ]
```

Now. we fit a spline model. Due to limitation,, we only show the prediction performance and ignore the plot.
Two breaking point are recognized in the first graph

```{r}
knot_age <- c(5,25)
knot_height <- c(70,137)
spline_model <- lm (weight ~ bSpline(height, knots = knot_height)+bSpline(age, knots = knot_age),
                       data = train_data)
# Make predictions
spline_predictions <- spline_model %>% predict(test_data)
# Model performance
spline_model_performance <- data.frame(
  RMSE = RMSE(spline_predictions, test_data$weight),
  R2 = R2(spline_predictions, test_data$weight)
)
summary(spline_model)
spline_model_performance
```


Finally a GAM model will be fitted to our data
```{r}
gam_model <- gam(weight~s(height) + s(age), data = train_data)
predictions_gam <- gam_model %>% predict(test_data)
# Model performance
gam_model_performance <- data.frame(
RMSE = RMSE(predictions_gam, test_data$weight),
R2 = R2(predictions_gam, test_data$weight)
)
summary(gam_model)
gam_model_performance
draw(gam_model, residuals = TRUE)

```


The above graph demonestrare that Our model predicts the height properly but, can not fit well the age parameter.
GAM and Spline model approximately give the same result.GAM model presents less RMSE than Spline model(about0.04).
both give the r2=0.94. If we divide the data into more knots in Spline model, we may get the same result.
